---
title: "Statistics 101B Project -- Caffeine & Attention"
author: "Group 4: Nicholas Cassol-Pawson, Max Chalekson, Romy Gou, Kanzah Jamil, Oliver Siu, Emma Vidal, Jesper White"
date: "June 9, 2024"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    code_folding: show
  pdf_document:
    toc: false
    toc_depth: 4
    keep_tex: yes
---

## 1. Sourcing

```{r source, message = FALSE}
library(tidyverse)
library(DescTools)
library(car)
library(pwr)
island_caffeine <- read_csv("island_caffeine.csv")
```


## 2. Design of the Experiment

### a) Design

```{r latin square}
latin <- function(n, random = FALSE){
  # generates a Latin Square of order n
  x <- matrix(LETTERS[1:n], n, n)
  for (j in 2:n) {
    x[, j] <- x[c(j:n, 1:(j - 1)), j]
  }
  if (random) {
    x <- x[sample(n), ]
    x <- x[, sample(n)]
  }
  x
}

latin(3)
set.seed(4)
latin(3, random = TRUE)
```


### b) Sampling

```{r sampling}
# island names
ironbark_names <- c("Hofn", "Vardo", "Helvig",
                    "Bjurholm", "Blonduos", "Helluland")
providence_names <- c("Hayarano", "Akkeshi", "Reading",
                      "Nelson", "Arcadia", "Kiyobico",
                      "Takazaki", "Shinobi", "Biruwa")
bonne_names <- c("Nidoma", "Colmar", "Riroua",
                 "Pauma", "Talu", "Valais",
                 "Kinsale", "Mahuti", "Vaiku",
                 "Eden", "Maeva", "Gordes")

# town house counts
ironbark_house <- c(937, 596, 483,
                    434, 431, 387)
providence_house <- c(521, 461, 714,
                      318, 1557, 520,
                      416, 358, 451)
bonne_house <- c(640, 2037, 462,
                 399, 483, 361,
                 429, 1017, 400,
                 523, 457, 403)

island_rng <- function(counts, town_names) {
  # generates and randomizes town and house number combinations
  town <- rep(town_names, counts)
  house <- numeric(0)
  for (i in counts) {
    house <- c(house, seq_len(i))
  }
  combo <- tibble(town, house)
  combo[sample(nrow(combo), nrow(combo)), ]
}

# generate rng objects
set.seed(4)
ironbark_rng <- island_rng(ironbark_house, ironbark_names)
providence_rng <- island_rng(providence_house, providence_names)
bonne_rng <- island_rng(bonne_house, bonne_names)

# index rng objects as necessary for viewing
ironbark_rng
providence_rng
bonne_rng
```

### c) Experiment Data

#### i) Loading Data Into Workspace

```{r load data}
# treatment, blocks, replicates
caffeine <- island_caffeine$caffeine
age <- island_caffeine$age_range
island <- island_caffeine$island
user <- island_caffeine$replicate

# response
timed <- island_caffeine$timed_diff

# ANOVA
timed_aov <- aov(timed ~
                   factor(caffeine) +
                   factor(age) +
                   factor(island) +
                   factor(user))
```

#### ii) Initial Plots

```{r basic plots, cache = TRUE}
age_index <- as.integer(factor(age))
island_index <- as.integer(factor(island))
user_index <- as.integer(factor(user))
caf_col <- c("blue", "orange", "red")

par(mfrow = c(2, 2))
plot(timed ~ factor(caffeine))
plot(timed ~ caffeine, main = "age",
     col = caf_col[age_index], pch = age_index)
plot(timed ~ caffeine, main = "island",
     col = caf_col[island_index], pch = island_index)
plot(timed ~ caffeine, main = "user",
     col = rainbow(7)[user_index], pch = user_index)

par(mfrow = c(1, 2))
plot(timed ~ factor(age))
plot(timed ~ factor(island))

par(mfrow = c(1, 1))
plot(timed ~ factor(user))
```

### d) Sample Size and Power




```{r power of test}
mean0 <- mean(timed[caffeine == 0])
mean1 <- mean(timed[caffeine == 100])
mean2 <- mean(timed[caffeine == 200])

caffeine_range <- range(c(mean0, mean1, mean2))
caffeine_range

d <- abs(diff(caffeine_range))
mse <- summary(timed_aov)[[1]][["Mean Sq"]][5]
caf_sd <- sqrt(mse)
f <- d / caf_sd

caffeine_min <- pwr.anova.test(k = 3, f = f, sig.level = 0.05, power = 0.8)
caffeine_min
caffeine_pwr <- pwr.anova.test(k = 3, n = 21, f = f, sig.level = 0.05)
caffeine_pwr

caf_n <- caffeine_min$n
caf_pwr <- caffeine_pwr$power
```

#### i) Minimum and Maximum means

The minimum and maximum means of the treatment group are $`r caffeine_range`$.

#### ii) Standard Deviation

A preliminary estimate of $\sigma^2$ is obtained from $MS_E$ which is $\hat\sigma^2 = `r mse`$. Therefore, we have that $\sigma = `r caf_sd`$.

#### iii) Justifying Sample Size

From `pwr.anova.test()`, $n = `r caf_n`$, so for 0.8 test power, $n=34$ replicates are required which is a sample size of $N=102$.

#### vi) Reporting Test Power

Since there are $k=3$ treatment groups and $n=21$ observations per group,
we have that the power of the test is $`r caf_pwr`$.

## 3. Results and interpretation

### a) Summary

```{r summary}
summary(timed_aov)
```

### b) Diagnostic Plots

```{r diagnostic, cache=TRUE}
par(mfrow = c(1, 2))
plot(timed_aov, which = c(1, 2))
```

### c) Post-hoc Analysis

```{r post-hoc}
TukeyHSD(timed_aov)
```
